\documentclass[11pt,oneside]{amsart}
\usepackage[margin=1in]{geometry}
\usepackage{amssymb,parskip,mathtools,microtype}
\usepackage[shortlabels]{enumitem}

\theoremstyle{definition}
\newtheorem{problem}{Problem}

\newcommand{\bC}{\mathbb{C}}
\newcommand{\bF}{\mathbb{F}}
\newcommand{\bQ}{\mathbb{Q}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\bZ}{\mathbb{Z}}
\newcommand*\colvec[1]{\begin{psmallmatrix}#1\end{psmallmatrix}}
\newcommand*\dcolvec[1]{\begin{pmatrix}#1\end{pmatrix}}
\newcommand{\extp}{\mathchoice{{\textstyle\bigwedge}}%
    {{\bigwedge}}%
    {{\textstyle\wedge}}%
    {{\scriptstyle\wedge}}}

\DeclareMathOperator{\Span}{span}
\let\Re\relax
\DeclareMathOperator{\Re}{Re}
\let\Im\relax
\DeclareMathOperator{\Im}{Im}

\title{MATH2211 Spring 2022\\
Exam 2}
\author{Wednesday, April 20 2022}

\begin{document}
    \maketitle

    Name: \underline{\hspace{6cm}}

    This exam is open notes. There are 50 points total in this exam.

    \begin{problem}
        Let $A=\begin{pmatrix}
            -1&2\\
            3&-1
        \end{pmatrix}$.
        \begin{enumerate}[(a)]
            \item (5 points) Find an elementary matrix $E$ such that $EA$ is upper triangular.
            \vfill
            \item (5 points) Give a basis of $\extp^2\bR^2$ and write the matrix of the linear operator
            \[\extp^2 A\colon \extp^2 \bR^2\to \extp^2\bR^2\]
            in this basis.
            \vfill
        \end{enumerate}
    \end{problem}
    \newpage

    \begin{problem}
       (10 points) The \emph{right shift} operator $T\colon \bR^\infty\to\bR^\infty$ is defined by
       \[T(a_0,a_1,a_2,\dots)=(0,a_0,a_1,a_2,\dots).\]
       What are the eigenvalues and eigenvectors of $T$?
    \end{problem}
    \newpage

    \begin{problem}
        (10 points) Prove that if $T\colon V\to V$ is diagonalizable, then $T^2+T+I_V$ is also diagonalizable.

        Hint: Eigenvectors.
    \end{problem}
    \vfill
    \newpage

    \begin{problem}
        Let $A$ be a $50\times 100$ real matrix whose kernel is 74-dimensional.
       \begin{enumerate}[(a)]
           \item (5 points) What is the dimension of the space of row vectors $y$ such that $yA=0$?
           \vfill
           \item (5 points) Suppose you know that the vector $\colvec{1\\1\\\vdots\\1}\in \bR^{100}$ is in the kernel of $A$. This implies that every row vector $(y_1\ y_2\ \cdots\ y_{100})$ in the image of $^tA$ satisfies some condition. What is this condition? Prove your answer.
           \vfill
       \end{enumerate}
    \end{problem}
    \newpage

    \begin{problem}
        (10 points) Let $M=\begin{pmatrix}
            1&3\\1&1
        \end{pmatrix}$.
        For $n\geq 0$, let $a_n$ be the top left entry of $M^n$. Prove that
        \[\lim_{n\to\infty}\frac{a_n}{(1+\sqrt3)^n}=\frac 12.\]
    \end{problem}
    \vfill
\end{document}
