\documentclass[11pt,oneside]{amsart}
\usepackage{geometry}
\usepackage{amssymb,parskip,mathtools,microtype}
\usepackage[shortlabels]{enumitem}

\theoremstyle{definition}
\newtheorem{problem}{Problem}

\newcommand{\bC}{\mathbb{C}}
\newcommand{\bF}{\mathbb{F}}
\newcommand{\bQ}{\mathbb{Q}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\bZ}{\mathbb{Z}}
\newcommand*\colvec[1]{\begin{psmallmatrix}#1\end{psmallmatrix}}
\newcommand*\dcolvec[1]{\begin{pmatrix}#1\end{pmatrix}}

\DeclareMathOperator{\Span}{span}
\let\Re\relax
\DeclareMathOperator{\Re}{Re}
\let\Im\relax
\DeclareMathOperator{\Im}{Im}

\title{MATH2211 Spring 2022\\
Problem Set 4}
\author{Due Wednesday, February 23 2022 at 11:59 pm}

\begin{document}
    \maketitle

    Useful reading for this problem set: Axler pages 20-23 (Sums of subspaces), Axler page 47 (Dimension of a sum and proof)

    \begin{problem}
        For each $c\in\bR$, determine the dimension of
        \[U=\Span\left( \dcolvec{2\\3\\1},\dcolvec{1\\-1\\2},\dcolvec{7\\3\\c} \right).\]
    \end{problem}

    \begin{problem}
        Suppose that $U$ and $W$ are 4-dimensional $\bC$-subspaces of $\bC^6$. Show that one can find two vectors in $U\cap W$, neither of which is a scalar multiple of the other.\footnote{Hint: the existence of such two vectors is precisely equivalent to a simple dimension condition on $U\cap W$. Can you see it?}
    \end{problem}

    \begin{problem}
        \leavevmode\begin{enumerate}[(a)]
            \item Find a basis for the subspace
            \[U=\{(x_1,x_2,x_3,x_4)\in\bR^4:x_1+x_2=0\text{ and }x_3-x_4=0\}\]
            and extend it to a basis of $\bR^4$.

            \item Let $P_5(\bR)$ be the vector space of all polynomials of degree at most 5 with real coefficients. Find a basis for the subspace
            \[U=\{f(x)\in P_5(\bR):f(1)=0\text{ and }f(2)=0\}\]
            and extend it to a basis of $P_5(\bR)$.
        \end{enumerate}
    \end{problem}

    \begin{problem}
        Given subspaces $U_1,U_2\subseteq V$, prove that the following properties are equivalent.
        \begin{itemize}
            \item $V=U_1+U_2$ and $\{0\}=U_1\cap U_2$.
            \item Every vector $v\in V$ can be written in a \textbf{unique} way as $v=u_1+u_2$ with $u_1\in U_1$ and $u_2\in U_2$.
        \end{itemize}
        When these hold, we say that $V$ is the \emph{direct sum} of $U_1$ and $U_2$, and write $V=U_1\oplus U_2$.
    \end{problem}

    \begin{problem}
        Prove or give a counterexample: if $f\colon A\to B$ and $g\colon B\to C$ are invertible functions, then $g\circ f\colon A\to C$ is also invertible.
    \end{problem}
    
    \begin{problem}
        Suppose $T\colon V\to W$ is a linear map, and $v_1,\dots,v_n\in V$. For each statement, give a proof or a counterexample.
        \begin{enumerate}[(a)]
            \item If $T$ is injective and $v_1,\dots,v_n$ are linearly independent, then $T(v_1),\dots,T(v_n)$ are linearly independent.
            \item If $T(v_1),\dots,T(v_n)$ are linearly independent, then $v_1,\dots,v_n$ are linearly independent.
            \item If $T$ is surjective and $v_1,\dots,v_n$ span $V$, then $T(v_1),\dots,T(v_n)$ span $W$.
            \item If $T(v_1),\dots,T(v_n)$ span $W$, then $v_1,\dots,v_n$ span $V$.
        \end{enumerate}
    \end{problem}
\end{document}
