\documentclass[11pt,oneside]{amsart}
\usepackage{geometry}
\usepackage{amssymb,parskip,mathtools,microtype}
\usepackage[shortlabels]{enumitem}

\theoremstyle{definition}
\newtheorem{problem}{Problem}

\newcommand{\bC}{\mathbb{C}}
\newcommand{\bF}{\mathbb{F}}
\newcommand{\bQ}{\mathbb{Q}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\bZ}{\mathbb{Z}}
\newcommand*\colvec[1]{\begin{psmallmatrix}#1\end{psmallmatrix}}
\newcommand*\dcolvec[1]{\begin{pmatrix}#1\end{pmatrix}}

\DeclareMathOperator{\Span}{span}
\let\Re\relax
\DeclareMathOperator{\Re}{Re}
\let\Im\relax
\DeclareMathOperator{\Im}{Im}
\DeclareMathOperator{\im}{im}

\title{MATH2211 Spring 2022\\
Problem Set 6}
\author{Due Wednesday, March 23, 2022 at 11:59 pm}

\begin{document}
    \maketitle
    
    \begin{problem}
        Let $T\colon V\to V$ be a linear map from a vector space $V$ to itself. Let $\mathcal B=(v_1,\dots,v_n)$ be an ordered basis of $V$ and let $A\in M_n(F)$ be the matrix of $V$ with respect to $\mathcal B$. Let $\mathcal C=(v_1',\dots,v_n')$ be another ordered basis of $V$ and let $A'\in M_n(F)$ be the matrix of $T$ with respect to $\mathcal C$. Prove that there exists an invertible matrix $B\in M_n(F)$ such that $A'=BAB^{-1}$.
    \end{problem}

    \begin{problem}
        Factor $A=\begin{pmatrix}2&1&2\\0&1&0\\2&0&1\end{pmatrix}$ as a product of elementary matrices.

        Hint: One way to approach this is to try to turn $A$ into the identity matrix via elementary matrices, then see how what you've done is useful.
    \end{problem}

    \begin{problem}
        \leavevmode\begin{enumerate}[(a)]
            \item In this problem we show that a one-sided inverse of a square matrix is a two-sided inverse. More precisely, let $A,B\in M_n(F)$. Show that
            \[AB=I_n\iff BA=I_n.\]
            Hint: For the forward direction, assume $AB=I_n$ and first show that $A\colon F^n\to F^n$ is surjective. Similar hint applies to the other direction.

            \item In this problem we show that inverses are unique. More precisely, suppose $A,B,C\in M_n(F)$ such that $AB=I_n$ and $AC=I_n$. Prove that $B=C$.
            
            Hint: In this proof, you are not allowed to multiply both sides by $A^{-1}$, because doing that presupposes that $A^{-1}$ is unique! Instead, start by proving that $A$ is surjective.
        \end{enumerate}
    \end{problem}

    \begin{problem}
        Let $A\in M_n(\bR)$ be an invertible matrix with integer entries, such that $A^{-1}$ also has integer entries. Prove that $\det A=\pm 1$.
    \end{problem}

    \newpage

    \begin{problem}
        Compute:
        \begin{enumerate}[(a)]
            \item $\det\begin{pmatrix}
                0&1&2\\2&6&-1\\3&0&4
            \end{pmatrix}$.
            \item $\det\begin{pmatrix}
                2&0&2&-4\\12&6&6&1\\0&-1&4&5\\3&2&3&2
            \end{pmatrix}$.
        \end{enumerate}
    \end{problem}

    \begin{problem}
        Directly from the definition of $\det\begin{pmatrix}
            a&b\\c&d
        \end{pmatrix}=ad-bc$ for the $2\times 2$ determinant, prove the following:
        \begin{enumerate}[(a)]
            \item $\det(AB)=\det(A)\det(B)$ for all $A,B\in M_2(F)$.
            \item $A$ is invertible if and only if $\det(A)\neq 0$, in which case $A^{-1}=\frac 1{\det(A)}\begin{pmatrix}
                d&-b\\-c&a
            \end{pmatrix}$.
            \item The function $\det\colon \bR^2\times\bR^2\to\bR; (v,w)\mapsto \det\begin{pmatrix}
                v&w
            \end{pmatrix}$ (the determinant of the matrix whose columns are $v$ and $w$) is \emph{bilinear}, meaning that for all $\alpha,\beta\in F$ and $v_1,v_2,v,w_1,w_2,w\in\bR^2$, we have
            \begin{enumerate}[(i)]
                \item $\det(\alpha v_1+\beta v_2,w)=\alpha\det(v_1,w)+\beta\det(v_2,w)$,
                \item $\det(v,\alpha w_1+\beta w_2)=\alpha\det(v,w_1)+\beta\det(v,w_2)$.
            \end{enumerate}
        \end{enumerate}
    \end{problem}
\end{document}
